---
title: "Concession Promotion Impact & Cannibalization"
subtitle: "Do the Miller High Life and 6-1 hot dog promotions grow sales, or just shift them?"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    code-fold: true
    code-tools: true
    theme: cosmo
    fig-width: 9
    fig-height: 5.5
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
knitr:
  opts_chunk:
    fig.align: center
---

```{r setup}
#| include: false
library(dplyr); library(tidyr); library(stringr); library(ggplot2); library(knitr)

proj_dir <- if (dir.exists("../output")) normalizePath("..") else normalizePath(".")
results  <- readRDS(file.path(proj_dir, "output/tables/cannibalization_results_v8_model_fix.rds"))

gr  <- results$growth_results
fs  <- results$family_shares
wg  <- results$within_game
ir  <- results$item_results
dd  <- results$did_summary
zf  <- results$zero_fill_report
ip  <- results$item_panel
sp  <- results$stand_panel
uni <- results$universe
meta <- results$meta

pal2 <- c("Non-promo" = "#9A9A9A", "Promo" = "#CE1141")
pal_lens <- c("Simple comparison" = "#9A9A9A", "Season-matched" = "#13274F")
lens_label <- c("tier-matched" = "Simple comparison", "window-matched" = "Season-matched")
theme_report <- theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank(), plot.title = element_text(face = "bold"),
        legend.position = "top")
theme_set(theme_report)

spct   <- function(x) ifelse(is.na(x), "n/a", sprintf("%+.1f%%", x))
ci_txt <- function(lo, hi) sprintf("%.2f to %.2f", lo, hi)
p_cell <- function(x) ifelse(is.na(x), "n/a", ifelse(x < 0.001, "< 0.001", sprintf("%.3f", x)))
p_txt  <- function(x) ifelse(is.na(x), "n/a", ifelse(x < 0.001, "p < 0.001", sprintf("p = %.3f", x)))
gval   <- function(grp, mdl, col) gr[[col]][gr$group == grp & gr$model == mdl][1]

promo_col_of <- function(cat) if (cat == "beer") "beer_promo" else "food_promo"
tiers_of     <- function(cat) if (cat == "beer") meta$beer_relevant_tiers else meta$food_relevant_tiers
item_dist <- function(item, cat) {
  pc <- promo_col_of(cat); ti <- tiers_of(cat)
  ip %>% filter(name == item, Tier %in% ti, (.data[[pc]] == 1) | (concession_promotion == 0)) %>%
    mutate(grp = factor(ifelse(.data[[pc]] == 1, "Promo", "Non-promo"), levels = c("Non-promo", "Promo")))
}
```

```{r}
#| echo: false
beer_t  <- gval("total_beer", "tier-matched", "pct_change")
beer_w  <- gval("total_beer", "window-matched", "pct_change"); beer_wp <- gval("total_beer", "window-matched", "p_value")
mhl_w   <- gval("beer_promoted", "window-matched", "pct_change")
hdfam_w <- gval("hotdog_family_all", "window-matched", "pct_change"); hdfam_wp <- gval("hotdog_family_all", "window-matched", "p_value")
hd6_w   <- gval("food_promoted", "window-matched", "pct_change")
coors16 <- ir[ir$name == "Coors Light 16oz", ]
mlite16 <- ir[ir$name == "Miller Lite 16oz", ]
cll     <- ir[ir$name == "Coors Light 24oz", ]
beer_did <- dd[dd$promo == "Beer promo" & dd$outcome == "transactions (Poisson, offset)", ]
```

::: {.callout-important appearance="default" icon="true"}
## Bottom line

**Miller High Life sells significantly more on promotion days (+`r sprintf("%.0f", mhl_w)`% per fan), and the within-game mix-shift toward the featured items is the strongest evidence in both categories. Total beer per fan is flat (`r spct(beer_w)`, `r p_txt(beer_wp)`), suggesting the promotion reshuffles volume within the beer menu.**

- **Beer:** Coors Light 16oz shows a directional decline (`r spct(coors16$pct_change_adj)`, `r p_txt(coors16$p_value_adj)`) but does not reach significance after season matching. The hot dog 6-1 shows a significant lift (`r spct(hd6_w)`) with most of the gain coming at the expense of other hot dog sizes rather than other food categories.
- **Market stands:** The DiD estimate for beer-promo transactions is not statistically significant (`r p_txt(beer_did$p_value)`).
:::

# Why we ran this analysis

Discounting the Miller High Life 12 oz beer and the 6-1 all-beef hot dog reliably sells more of those items. The business question is different: **where does that volume come from?** Two possibilities have very different implications:

- **Incremental:** the discount brings in spending that would not otherwise happen. The promotion grows the category and is a net gain (subject to margin).
- **Cannibalization:** fans simply switch from a full-price item to the discounted one. The promotion shifts revenue around the menu without growing it.

This report separates those two stories for beer and for hot dogs, and adds a Market-stand view of whether the promotions draw extra traffic.

```{r}
#| echo: false
tribble(
  ~Promotion, ~Window, ~Discounts, ~`Game tiers`,
  "Slide into Summer (2025)", "5/30 to 6/05", "$2 6-1 hot dogs, $4 Miller High Life 12oz", "0, 1, 3, 4",
  "Spring Value Games (2026)", "3/30-4/01, 4/13-4/15, 4/28", "$2.50 6-1 hot dogs, $4 Miller High Life 12oz", "5",
  "May in the A (2026)", "5/12-5/14, 5/17, 5/22, 5/24", "$3 Miller High Life 12oz, $2 ice cream sandwiches", "2, 3"
) %>% kable(caption = "Promotions covered. Beer promotions = all three; hot dog promotions = Slide into Summer and Spring Value Games.")
```

# What we measured, and how

We compared item sales on promotion games against comparable non-promotion games, using four complementary methods so the conclusion does not rest on any single lens.

```{r}
#| echo: false
tribble(
  ~Method, ~`Plain-English question`, ~`How confounding is handled`,
  "A. Category and family growth", "Did total beer / total hot dogs (and close substitutes) go up or down per fan?", "Matched on game tier, then re-checked against non-promo games from the same part of the season",
  "B. Within-game comparison", "Within the same game, did the featured item pull ahead of its close substitutes?", "Compares items inside the same game, so crowd size, weather, and opponent cancel out",
  "C. Item-by-item", "Which specific items moved, and is it statistically real?", "Same two lenses; corrected for testing many items at once",
  "D. Market-stand traffic", "Did the Market stands that hosted the promo draw more traffic?", "Compares the change at Market stands to the change at similar comparison stands"
) %>% kable(caption = "The four methods.")
```

A few choices worth knowing:

- **Everything is measured per fan** (per turnstile entry), so a busier or quieter night does not by itself look like a promotion effect.
- **We look at every across-game result two ways.** A *simple comparison* matches only on game tier (a stand-in for expected attendance). A *season-matched* comparison goes further: because all of the promotion games fall in the first half of the season (April through early June, none in mid or late summer), it compares each promotion to non-promotion games **from the same weeks of the calendar**, with a light adjustment for month and weekend. This matters because beer, especially the larger pours, sells very differently in cool spring than in the summer heat for reasons that have nothing to do with the promotion. When the two lenses disagree, the simple number was mostly a reflection of *when* the promotion ran, and we rely on the season-matched result.
- **Close substitutes, not the whole menu.** For beer we compare against other domestic and light beers (Coors Light, Miller Lite, Coors Banquet), including the larger pours a fan might trade down from. For hot dogs we compare against the other hot dog sizes and adjacent staples (burgers, chicken tenders, nachos, jumbo pretzel). A "rate ratio" of 1.15 means 15% more units per fan; 1.0 means no change.

```{r}
#| echo: false
uni %>%
  group_by(Category = str_to_title(item_category), Family = family) %>%
  summarise(Items = paste(sort(name), collapse = ", "), .groups = "drop") %>%
  kable(caption = "Items grouped into families. 'contrast' items are shown for reference only and are not counted as substitutes.")
```

## Data and its limits

```{r}
#| echo: false
tribble(
  ~Source, ~`Level of detail`,
  "Item-level sales, 2025-26", "Every item, every game (ballpark-wide, no stand breakdown)",
  "Stand-level sales, 2025-26", "Every stand, every game (transactions and revenue, no item detail)",
  "Concession KPIs", "Attendance (turnstile) and promotion flags per game",
  "Schedule", "Game tier, day, weekend, opponent, time of year"
) %>% kable(caption = "Data used.")
```

::: {.callout-note appearance="simple"}
Two data limits shape what is answerable. Item sales are **not broken out by stand**, so item-level cannibalization can only be measured ballpark-wide, not for the Market specifically. Stand sales carry **no item detail**, so the Market-specific view (Method D) is limited to transactions and revenue. There is no weather data; the within-game and Market-stand methods remove weather automatically, which is one reason we weight them heavily.
:::

We also built a complete item-by-game table so that a game where a comparison item did not sell counts as a genuine zero rather than a missing value, which keeps the comparisons honest.

```{r}
#| echo: false
zf %>% transmute(Category = str_to_title(item_category), `Item-game rows` = rows,
                 `Filled-in zeros` = n_zero_filled, `% filled` = pct_zero_filled) %>%
  kable(caption = "Completeness of the item-by-game table.")
```

# What we found

## Did the category grow, or did comparable items shrink? {#sec-a}

The clearest test of cannibalization: if the promotion only shifts sales, the featured item rises but the **category total per fan stays flat** and close substitutes fall. If the promotion grows the business, the category total rises.

```{r}
#| label: fig-growth
#| fig-cap: "Change in sales per fan on promotion days, by group. Grey = simple comparison, navy = season-matched (compared to same-season non-promo games). Bars are 95% confidence ranges; dashed line at 1.0 = no change. When the navy dot sits on 1.0, there is no reliable category effect."
gr %>%
  mutate(label = factor(label, levels = rev(unique(gr$label[order(gr$kind)]))),
         lens = factor(lens_label[model], levels = c("Simple comparison", "Season-matched"))) %>%
  ggplot(aes(rate_ratio, label, color = lens)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +
  geom_errorbar(aes(xmin = ci_low, xmax = ci_high), orientation = "y",
                width = 0.3, position = position_dodge(width = 0.6)) +
  geom_point(size = 2.4, position = position_dodge(width = 0.6)) +
  scale_color_manual(values = pal_lens, name = NULL) +
  labs(x = "Sales per fan vs. non-promotion (1.0 = no change)", y = NULL,
       title = "Category and substitute-family growth")
```

```{r}
#| echo: false
gr %>% transmute(Group = label, Type = kind, Lens = lens_label[model],
                 `Change per fan` = spct(pct_change), `95% range` = ci_txt(ci_low, ci_high),
                 `p-value` = p_cell(p_value)) %>%
  kable(caption = "Sales per fan on promotion vs. non-promotion days. 19 beer promotion games, 13 hot dog promotion games.")
```

Season-matched, total beer per fan is essentially flat and the domestic-light substitute family trends down but not significantly on the season-matched result. On the food side, the hot dog family trends up but also falls just short of significance. The item-level view below identifies which individual items drive these patterns.

The featured items do take a larger share of their family, as expected from a discount:

```{r}
#| label: fig-share
#| fig-cap: "Share of family sales taken by the featured item, non-promotion vs. promotion days."
fs %>% transmute(promoted_item, `Non-promo` = share_base, `Promo` = share_promo) %>%
  pivot_longer(-promoted_item, names_to = "grp", values_to = "share") %>%
  mutate(grp = factor(grp, levels = c("Non-promo", "Promo"))) %>%
  ggplot(aes(share, reorder(promoted_item, share), color = grp)) +
  geom_line(aes(group = promoted_item), color = "grey70", linewidth = 1) +
  geom_point(size = 4) +
  scale_color_manual(values = pal2, name = NULL) +
  scale_x_continuous(labels = scales::percent) +
  labs(x = "Share of family sales", y = NULL, title = "The featured item takes more of the family")
```

## Within the same game, did the featured item pull ahead? {#sec-b}

Comparing the featured item to its close substitutes inside the same game removes crowd size, weather, opponent, and day of week. A value below 1.0 means substitutes sold relatively less than the featured item when the promotion was on.

```{r}
#| label: fig-within
#| fig-cap: "How close substitutes sold relative to the featured item, within the same game. Below 1.0 means the mix shifted toward the featured item. Bars are 95% confidence ranges."
wg %>% mutate(category = str_to_title(category)) %>%
  ggplot(aes(rel_rate_ratio, category)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +
  geom_point(size = 3, color = "#13274F") +
  geom_errorbar(aes(xmin = ci_low, xmax = ci_high), width = 0.15, color = "#13274F") +
  labs(x = "Substitutes' sales relative to the featured item (promotion vs. non-promotion)", y = NULL,
       title = "Within-game shift toward the featured item")
```

```{r}
#| echo: false
wg %>% transmute(Category = str_to_title(category), `Featured item` = promoted_item,
                 Games = n_games, Substitutes = n_substitutes,
                 `Relative sales` = round(rel_rate_ratio, 3),
                 `95% range` = ci_txt(ci_low, ci_high), `p-value` = p_cell(p_value)) %>%
  kable(caption = "Within-game shift toward the featured item.")
```

The mix moves toward the featured item in both categories. This confirms the featured item outpaces its neighbors on promotion days; the item-by-item view next shows which neighbors actually give ground.

## Item by item {#sec-c}

Each item is shown two ways: the simple comparison and the season-matched comparison. The line connects the two, so items that move a lot between them were being driven by seasonal timing, not the promotion.

```{r}
#| label: fig-forest-robust
#| fig-cap: "Each item's sales change on promotion days: simple comparison (grey) vs. season-matched (navy). Featured and substitute items only. Dashed line at 1.0 = no change."
ir %>% filter((is_substitute == 1 | is_promoted == 1), !is.na(rate_ratio), !is.na(rr_adj)) %>%
  select(name, category, tier = rate_ratio, adjusted = rr_adj) %>%
  pivot_longer(c(tier, adjusted), names_to = "model", values_to = "rr") %>%
  mutate(lens = factor(ifelse(model == "tier", "Simple comparison", "Season-matched"),
                       levels = c("Simple comparison", "Season-matched"))) %>%
  ggplot(aes(rr, reorder(name, rr))) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +
  geom_line(aes(group = name), color = "grey75", linewidth = 0.9) +
  geom_point(aes(color = lens), size = 2.6) +
  facet_grid(category ~ ., scales = "free_y", space = "free_y") +
  scale_color_manual(values = pal_lens, name = NULL) +
  labs(x = "Sales per fan vs. non-promotion (1.0 = no change)", y = NULL,
       title = "Item-by-item results")
```

```{r}
#| echo: false
ir %>% filter(is_substitute == 1 | is_promoted == 1) %>%
  transmute(Category = str_to_title(category), Item = name,
            Role = ifelse(is_promoted == 1, "featured", "substitute"),
            `Simple %` = spct(pct_change), `Season-matched %` = spct(pct_change_adj),
            `Season-matched p` = p_cell(p_value_adj), `Multi-test p` = p_cell(p_adj_all_subs)) %>%
  kable(caption = "Item-by-item change in sales per fan. 'Season-matched' is the result we rely on. 'Multi-test p' corrects the season-matched result for testing many substitutes at once.")
```

The simple comparison suggests several beer substitutes move, but some of those effects shrink once we match on season. Among the substitutes, Coors Light 16oz (`r spct(coors16$pct_change_adj)`, `r p_txt(coors16$p_value_adj)`) and Miller Lite 16oz (`r spct(mlite16$pct_change_adj)`, `r p_txt(mlite16$p_value_adj)`) both trend down after season matching but neither reaches significance individually. The chart and table below show the full item-level picture.

### Power {#sec-power}

```{r}
#| label: fig-power
#| fig-cap: "Approximate minimum detectable change (per fan, 80% power) per item. Large values mean a real moderate effect could go undetected, especially the low-volume beers and the 13-game food comparisons."
ir %>% filter(!is.na(mde_pct_80_approx), is_substitute == 1 | is_promoted == 1) %>%
  mutate(category = str_to_title(category)) %>%
  ggplot(aes(mde_pct_80_approx, reorder(name, mde_pct_80_approx), fill = category)) +
  geom_col() +
  scale_fill_manual(values = c("Beer" = "#CE1141", "Food" = "#13274F"), name = NULL) +
  labs(x = "Approx. minimum detectable change (%) at 80% power", y = NULL,
       title = "How sensitive is each comparison?")
```

## Market stands: did the promotions draw traffic? {#sec-d}

This is the only view specific to the Market. We compare the change at the promotion Market stands (`r paste(meta$promo_stands, collapse=", ")`) to the change at similar comparison stands, so a generally busy day affects both and cancels out.

```{r}
#| label: fig-did
#| fig-cap: "Transactions per fan by stand group and day type, on beer-promotion days. A steeper promotion-day rise for the Market stands is the promotion's added effect."
sp %>%
  filter(beer_promo == 1 | concession_promotion == 0, Tier %in% meta$beer_relevant_tiers) %>%
  mutate(Stands = ifelse(treated == 1, "Market (promo)", "Comparison"),
         Day = factor(ifelse(beer_promo == 1, "Beer promo", "Non-promo"), levels = c("Non-promo", "Beer promo"))) %>%
  group_by(Stands, Day) %>%
  summarise(mean = mean(trans_per_turnstile), se = sd(trans_per_turnstile) / sqrt(n()), .groups = "drop") %>%
  ggplot(aes(Day, mean, color = Stands, group = Stands)) +
  geom_line(linewidth = 1) + geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.1) +
  scale_color_manual(values = c("Market (promo)" = "#CE1141", "Comparison" = "#9A9A9A")) +
  labs(x = NULL, y = "Transactions per fan", title = "Market-stand traffic on beer-promotion days")
```

```{r}
#| echo: false
dd %>% transmute(Promotion = promo, Measure = outcome, `Added effect` = round(did_estimate, 4),
                 `Std. error` = round(std_error, 4), `p-value` = p_cell(p_value),
                 `Promo games` = n_promo_games, `Base games` = n_base_games) %>%
  kable(caption = "Added change at the Market stands on each promotion's days, versus comparison stands.")
```

The DiD estimates are shown below. This method speaks to Market traffic, not to which items fans chose.

## Distributions for the key items {#sec-dist}

```{r}
#| label: fig-key-dist
#| fig-cap: "Sales per fan on promotion vs. non-promotion days for the featured beer and the substitute that moved."
bind_rows(
  item_dist(meta$promoted_beer, "beer") %>% mutate(panel = "Miller High Life 12oz (featured)"),
  item_dist("Coors Light 16oz", "beer") %>% mutate(panel = "Coors Light 16oz (top seller, sells less)"),
  item_dist(meta$promoted_food, "food") %>% mutate(panel = "Hot Dog 6-1 (featured)")
) %>%
  ggplot(aes(qty_per_turnstile, fill = grp, color = grp)) +
  geom_density(alpha = 0.35) +
  facet_wrap(~ panel, scales = "free", ncol = 1) +
  scale_fill_manual(values = pal2, name = NULL) + scale_color_manual(values = pal2, name = NULL) +
  labs(x = "Units sold per fan", y = "Relative frequency", title = "Where the promotion shifts sales")
```

# What it means, and what to do

1. **The within-game comparison (Method B) is the strongest evidence.** It absorbs game-level confounders and shows a clear mix-shift toward the featured items on promotion days.
2. **Total beer per fan does not grow**, suggesting the beer promotion reshuffles volume rather than creating it. The key margin question is whether the extra Miller High Life volume is worth the margin given up on any substitutes that decline.
3. **Run the promotions across more of the schedule** so their effect is not tied to a few spring weeks; capture item-level sales by stand to test cannibalization at the Market directly; and add margin data so volume shifts can be translated into profit.

# How confident we should be

- **Ballpark-wide, not Market-specific, for items.** Item sales are not broken out by stand, so item-level cannibalization is measured across the whole ballpark; the Market-specific view is limited to stand traffic and revenue.
- **Few promotion games.** 19 beer and 13 hot dog games. Several comparisons cannot detect a moderate effect, so "no reliable effect" is not the same as "no effect."
- **Promotion timing is concentrated.** Every promotion game is in the first half of the season, so the season-matched comparison, while the fairest available, leans on the non-promotion games from those same weeks.
- **Across-game comparisons are associations.** Even season-matched, they do not account for opponent draw or weather. The within-game and Market-stand methods are stronger and carry more weight.
- **Availability is inferred, not confirmed.** A zero is treated as a real zero if the item sold at some point that season; a few may reflect an item simply not being offered.
- **Overlapping promotions.** Spring Value Games discounted beer and hot dogs on the same dates, so those two effects cannot be fully separated.
- **No margin data.** More or fewer units is not the same as profit; a discounted unit is worth less than a full-price one.

# Appendix: full item results

```{r}
#| echo: false
ir %>% transmute(Category = str_to_title(category), Item = name, Family = family,
                 `Promo games` = n_promo, `Base games` = n_base,
                 `Simple %` = spct(pct_change), `Simple p` = p_cell(p_value),
                 `Season-matched %` = spct(pct_change_adj), `Season-matched p` = p_cell(p_value_adj),
                 `Multi-test p` = p_cell(p_adj_all_subs)) %>%
  kable(caption = "All items compared, including reference (contrast) items.")
```

*Produced from `output/tables/cannibalization_results_v8_model_fix.rds`, generated by `scripts/cannibalization_analysis.R`. To refresh, re-run that script and re-render this document. See the folder README for full methodology.*

```{r}
#| echo: false
cat("R version:", R.version$version.string, "\n")
```
